Three distinct patterns of mental health response following accidents in mountain sports – a follow-up study of individuals treated at a tertiary trauma center

### 2023-10-11 paper/markdown/manuscript.Rmd
clustTools and caretExtra compatibility
Compatibility with the newest package versions and new tools

The cluster assignment models employing early predictors demonstrated a moderate-to-excellent prediction performance in the training subset of the study cohort 
(`r paste0((filter(early_class$overall_stats, dataset == 'train')$correct_rate %>% range %>% signif(2)) * 100, '%') %>% paste(collapse = ' - ')`% of correctly assigned observations, 
Cohen's $\kappa$: `r signif(range(filter(early_class$overall_stats, dataset == 'train')$kappa), 2) %>% paste(collapse = ' - ')`, Brier score: `r signif(range(filter(early_class$overall_stats, dataset == 'train')$bs), 2) %>% paste(collapse = ' - ')`). 
However, their accuracy in cross-validation and in the test subset was poor 
(test: `r paste0((filter(early_class$overall_stats, dataset == 'test')$correct_rate %>% range %>% signif(2)) * 100, '%') %>% paste(collapse = ' - ')`% of correctly assigned observations, 
Cohen's $\kappa$: `r signif(range(filter(early_class$overall_stats, dataset == 'test')$kappa), 2) %>% paste(collapse = ' - ')`, 
Brier score: `r signif(range(filter(early_class$overall_stats, dataset == 'test')$bs), 2) %>% paste(collapse = ' - ')`). 
Concerning the early identification of the vulnerable PTS cluster patients, the conditional random forest, random forest and support vector machine models were able to differentiate PTS cluster patients from the pooled neutral + PTG cluster subset with the best, at least 
`r min(signif(early_class$clust_stats %>% filter(dataset == 'test', clust_id == 'PTS') %>% top_n(n = 3, sens) %>% .$sens, 2) * 100)`% sensitivity in the test subset of the study cohort (__Figure \@ref(fig:fig-early-class)__, __Supplementary Table S12__). 
Pre-existing mental disorder or physical illness, hospitalization and surgery, injury severity, age, rescue mode, education, annual income, and traumatic events in the past were among the most important explanatory factors employed by these models (__Supplementary Figure S8__). 
Of note, inclusion of follow-up-related variables such as psychological support, physical health consequences or caution during sport activity in the machine learning models of the mental cluster assignment did not improve their accuracy in cross-validation or the test subset of the study cohort 
(test: `r paste0((filter(full_class$overall_stats, dataset == 'test')$correct_rate %>% range %>% signif(2)) * 100, '%') %>% paste(collapse = ' - ')`% of correctly assigned observations, 
Cohen's $\kappa$: `r signif(range(filter(full_class$overall_stats, dataset == 'test')$kappa), 2) %>% paste(collapse = ' - ')`, 
Brier score: `r signif(range(filter(full_class$overall_stats, dataset == 'test')$bs), 2) %>% paste(collapse = ' - ')`).

>>>>

The cluster assignment models employing early predictors demonstrated a moderate-to-excellent prediction performance in the training subset of the study cohort 
(`r paste0((filter(early_class$overall_stats, dataset == 'train')$correct_rate %>% range %>% signif(2)) * 100, '%') %>% paste(collapse = ' - ')`% of correctly assigned observations, 
Cohen's $\kappa$: `r signif(range(filter(early_class$overall_stats, dataset == 'train')$kappa), 2) %>% paste(collapse = ' - ')`, Brier score: `r signif(range(filter(early_class$overall_stats, dataset == 'train')$brier_score), 2) %>% paste(collapse = ' - ')`). 
However, their accuracy in cross-validation and in the test subset was poor 
(test: `r paste0((filter(early_class$overall_stats, dataset == 'test')$correct_rate %>% range %>% signif(2)) * 100, '%') %>% paste(collapse = ' - ')`% of correctly assigned observations, 
Cohen's $\kappa$: `r signif(range(filter(early_class$overall_stats, dataset == 'test')$kappa), 2) %>% paste(collapse = ' - ')`, 
Brier score: `r signif(range(filter(early_class$overall_stats, dataset == 'test')$brier_score), 2) %>% paste(collapse = ' - ')`). 
Concerning the early identification of the vulnerable PTS cluster patients, the conditional random forest, random forest and support vector machine models were able to differentiate PTS cluster patients from the pooled neutral + PTG cluster subset with the best, at least 
`r min(signif(early_class$clust_stats %>% filter(dataset == 'test', clust_id == 'PTS') %>% top_n(n = 3, sens) %>% .$sens, 2) * 100)`% sensitivity in the test subset of the study cohort (__Figure \@ref(fig:fig-early-class)__, __Supplementary Table S12__). 
Pre-existing mental disorder or physical illness, hospitalization and surgery, injury severity, age, rescue mode, education, annual income, and traumatic events in the past were among the most important explanatory factors employed by these models (__Supplementary Figure S8__). 
Of note, inclusion of follow-up-related variables such as psychological support, physical health consequences or caution during sport activity in the machine learning models of the mental cluster assignment did not improve their accuracy in cross-validation or the test subset of the study cohort 
(test: `r paste0((filter(full_class$overall_stats, dataset == 'test')$correct_rate %>% range %>% signif(2)) * 100, '%') %>% paste(collapse = ' - ')`% of correctly assigned observations, 
Cohen's $\kappa$: `r signif(range(filter(full_class$overall_stats, dataset == 'test')$kappa), 2) %>% paste(collapse = ' - ')`, 
Brier score: `r signif(range(filter(full_class$overall_stats, dataset == 'test')$brier_score), 2) %>% paste(collapse = ' - ')`).

### 2023-10-11 paper/markdown/supplementary_material.Rmd
clustTools and caretExtra compatibility
Compatibility with the newest package versions and new tools

$$BS = \frac{1}{N}\times\Sigma_{i = 1}^{N}\Sigma_{c = 1}^R(f_{ic} - o_{ic})^2$$
where $f_{tc}$ is the output probability of assignment of the i-th observation to the c-th cluster, 
$o_{ic}$ is the numeric-coded actual assignment of the i-th observation to the c-th cluster, 
$R$ is the total cluster number and 
$N$ is the total observation number. 
Brier skill scores [@Brier1950] were computed with the following formula:

$$BSS = 1 - \frac{BS_{model}}{BS_{random}}$$

where $BS_{model}$ is the Brier score of the model and 
$BS_{random}$ is the Brier score obtained for a purely random cluster assignment. 
Brier scores, Brier skill scores, sensitivity and specificity of the PTS cluster assignment were computed with in-house developed R functions. 
Performance statistics for the training subset, 10-fold cross-validation and test subset of the study cohort are listed in __Supplementary Table S\@ref(tab:tab-early-classifiers)__ and __Supplementary Table S\@ref(tab:tab-full-classifiers)__ for the classifiers employing the early and full predictor set, respectively.

@@

flextable::flextable(suppl_tables$early_classifiers) %>% 
  set_widths(c(2.3, 2.5, rep(2.3, 6))) %>% 
  merge_v(1) %>% 
  footnote(1, 1, 
           value = as_paragraph('RF: random forest; NNet: neural network with a single hidden layer; SVM/radial: support vector machines with radial kernel; RPart: recursive partitioning; SDA: shrinkage discriminant analysis; cForest: conditional random forest; Elastic Net: elastic net multinomial regression.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 2, 
           value = as_paragraph('CV: cross-validation'), 
           part = 'header',  
           ref_symbols = 'b') %>% 
  footnote(1, 6, 
           value = as_paragraph('Brier Skill Score comparing the given classifier with the purely random cluster assignment.'), 
           part = 'header', 
           ref_symbols = 'c') %>% 
  theme_vanilla

@@

flextable::flextable(suppl_tables$full_classifiers) %>% 
  set_widths(c(2.3, 2.5, rep(2.3, 6))) %>% 
  merge_v(1) %>% 
  footnote(1, 1, 
           value = as_paragraph('RF: random forest; NNet: neural network with a single hidden layer; SVM/radial: support vector machines with radial kernel; RPart: recursive partitioning; SDA: shrinkage discriminant analysis; cForest: conditional random forest; Elastic Net: elastic net multinomial regression.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 2, 
           value = as_paragraph('CV: cross-validation'), 
           part = 'header',  
           ref_symbols = 'b') %>% 
  footnote(1, 6, 
           value = as_paragraph('Brier Skill Score comparing the given classifier with the purely random cluster assignment.'), 
           part = 'header', 
           ref_symbols = 'c') %>% 
  theme_vanilla

>>>>

$$BS = \frac{1}{N}\times\Sigma_{i = 1}^{N}\Sigma_{c = 1}^R(f_{ic} - o_{ic})^2$$
where $f_{tc}$ is the output probability of assignment of the i-th observation to the c-th cluster, 
$o_{ic}$ is the numeric-coded actual assignment of the i-th observation to the c-th cluster, 
$R$ is the total cluster number and 
$N$ is the total observation number.

Performance statistics for the training subset, 10-fold cross-validation and test subset of the study cohort are listed in __Supplementary Table S\@ref(tab:tab-early-classifiers)__ and __Supplementary Table S\@ref(tab:tab-full-classifiers)__ for the classifiers employing the early and full predictor set, respectively.

@@

flextable::flextable(suppl_tables$early_classifiers) %>% 
  set_widths(c(2.3, 2.5, rep(2.3, 5))) %>% 
  merge_v(1) %>% 
  footnote(1, 1, 
           value = as_paragraph('RF: random forest; NNet: neural network with a single hidden layer; SVM/radial: support vector machines with radial kernel; RPart: recursive partitioning; SDA: shrinkage discriminant analysis; cForest: conditional random forest; Elastic Net: elastic net multinomial regression.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 2, 
           value = as_paragraph('CV: cross-validation'), 
           part = 'header',  
           ref_symbols = 'b') %>% 
  theme_vanilla

@@

flextable::flextable(suppl_tables$full_classifiers) %>% 
  set_widths(c(2.3, 2.5, rep(2.3, 5))) %>% 
  merge_v(1) %>% 
  footnote(1, 1, 
           value = as_paragraph('RF: random forest; NNet: neural network with a single hidden layer; SVM/radial: support vector machines with radial kernel; RPart: recursive partitioning; SDA: shrinkage discriminant analysis; cForest: conditional random forest; Elastic Net: elastic net multinomial regression.'), 
           part = 'header', 
           ref_symbols = 'a') %>% 
  footnote(1, 2, 
           value = as_paragraph('CV: cross-validation'), 
           part = 'header',  
           ref_symbols = 'b') %>% 
  theme_vanilla
